\subsection{Theoretical Limitations of Measurement}
\subsubsection{Range Measurement}
While considering radar limitations, it is necessary to mention range resolution and range ambiguity. Range resolution is a limiting ability to distinguish two targets based on their range difference. This is possible by measuring the time delay between transmitted and received waves. Range resolution depends on signal bandwidth $B$ according to the equation:
\[\Delta R = \frac{c}{2B},\]
where $c$ is the wave propagation speed.
Unambiguous range measurement using a system with a constant pulse repetition interval $T_\mathrm{PRI}$ is only possible when the actual target's range is less than
\[R_\mathrm{max} = c \cdot T_\mathrm{PRI}/2\].

\subsubsection{Velocity Measurement}
Except for range, radar can also measure a target's velocity thanks to the Doppler frequency shift. Velocity resolution is a limiting ability to distinguish two targets based on their velocity difference. It depends on the carrier frequency $f_c$, pulse repetition frequency $f_\mathrm{PRF}$, and observation time T according to the equation:
\[\Delta v = \] % uzupelnic
Taking advantage of the velocity resolution, it is possible to separate echoes originating from different parts of the human body performing some specific movements. Analysis of how coexisting velocity components change in time is a typical way for observation and classification of human behavior using radar.
Unambiguous velocity measurement using a system with constant pulse repetition interval is only possible when the actual target's velocity is less than
\[v_\mathrm{max} = \]  % uzupelnic
Otherwise, the echo wraps around the speed range $[-v_\mathrm{max}, v_\mathrm{max}]$.

\subsubsection{Phase Measurement}
It is worth noting that taking advantage of some assumptions, it is possible to measure the exact displacement waveform of the target based on the phase changes between consecutive pulses. If the reflection mostly originates from a single target, the signal's phase in the slow time corresponds to the relative displacement according to the equation:
\[\Phi(t) = ??? \cdot r(t).\] % uzupelnic

One of the problems is that the measured phase is wrapped within the interval $[0, 2\pi)$. For this reason, the raw phase waveform needs to be unwrapped to extract a valid displacement waveform. This causes the noise impact to be additionally problematic. 
%%% jakies zrodlo potwierdzajace
Either way, signal phase analysis is the most commonly used method to observe human breathing and heart rate.

\subsubsection{Wave Resistance to Obstacles}